---
title: "DS6306 Case Study 01"
author: "Ranjan Karki & Jason McDonald"
date: "12/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Exploratory Data Analysis of Breweries and Beer 
## Prepared for Brendan Whitworth, CEO, and Fernando Tennenbaum, CFO, of Anheuser Busch InBev  
Mr. Whitworth, Mr. Tennenbaum, thank you for the opportunity to work with Anheuser Busch InBev to analyze data on a set of craft beers and the breweries that make them.  We've received the datasets containing information on over 2400 craft beers made by over 550 breweries.   

We've begun to explore these datasets with the intention of answering the questions you previously provided.  I believe we've been able to extract insights that you will find valuable to Anheuser Busch InBev.    

With that, if you are ready, let's walk through our initial findings within the data.  


## Question 1: How many breweries are present in each state?
### To approach this question, we were able to read in the data contained in the list of Breweries and count the number in each state. As you can see, Colorado leads with 47 breweries while 4 different states, West Virginia, South Dakota, North Dakota, and Washington DC only have 1 brewery.
```{r "Question 1", message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
#Read in Breweries.csv and setup a list with the state and count of each state, then build a barplot
Breweries <- read.csv("Breweries.csv", header = TRUE , ",")
totals <- Breweries %>% count (State)
q1 <- ggplot(data=totals, aes(x = State,y=n, fill = n, width = 0.8)) + geom_bar(stat='identity', width=0.7) + scale_fill_gradient2(midpoint=mean(totals$n), low= '#971B1E', mid = '#E02124', high = '#FFF200', space='#E1E1E1') + geom_text(aes(x= State,n +1, label = paste(State, n, sep=' - ')), data = totals, position = position_dodge(width = 1), hjust = -0.1, size = 4) + ggtitle("Total Breweries in each State") + coord_flip() + theme(legend.title = element_text(colour='#331D0C')) + labs(fill="Count of Breweries", y = "Count of Breweries") + scale_y_continuous(limits = c(0,max(totals$n) *1.1))
```

### The below barplot displays the count of breweries per state. 
```{r "Question 1 Result", message = FALSE, warning = FALSE, fig.dim = c(11,9)}
q1
```

## Question 2: Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

```{r "Question 2", message = FALSE, warning = FALSE}
Beers <- read.csv("Beers.csv", header = TRUE , ",")
BeerN <- rename(Beers,"Brew_ID" = "Brewery_id")
tdf <- merge(BeerN,Breweries, by = "Brew_ID", na.rm = True) %>% rename(Beer = Name.x, Brewery = Name.y)

#View(tdf)
head(tdf,n=6)
tail (tdf,n=6)

###EXTRA
top_n(tdf,6,Brew_ID)

```

## Question 3: Address the missing values in each column.
### We found 1005 beers which didn't have an IBU value and 62 that didn't have an ABV.  As your questions pertain to these values and there is no method that we can use to approximate the missing values, removing these beers from the dataset was the only option.
```{r "Question 3", message = FALSE, warning = FALSE}

colSums(is.na(tdf))
  
### assigning to new data frame 
dtdf <- tdf[!is.na(tdf$ABV), ]
dtdf <- tdf[!is.na(tdf$IBU), ]


colSums(is.na(dtdf))

```

## Question 4: Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
### To identify the median ABV and IBU by state, we performed a calculation on the dataset, grouping by state and finding the mediam in each.  We then generated barplots as seen here.  

```{r "Question 4", message = FALSE, warning = FALSE, fig.dim = c(11,9)}

df1<- dtdf %>% group_by(State) %>% summarise(Median_ABV = median(ABV)) 
ggplot(data = df1, aes(x= State, y = Median_ABV * 100 ,fill = Median_ABV))+geom_bar(stat='identity', width=0.7) + ggtitle ("State v.Median_ABV") + scale_fill_gradient2(midpoint=mean(df1$Median_ABV), low= '#971B1E', mid = '#E02124', high = '#FFF200', space='#E1E1E1') + coord_flip() + geom_text(data = df1, aes(x = State, label = paste(State, paste(Median_ABV * 100, '%', sep = ''), sep=' - ')), position = position_dodge(width = 1), hjust = -0.1, size = 4) + scale_y_continuous(limits = c(0,max(df1$Median_ABV * 100) *1.1)) 
## question how can I do both at the same time 


df2<- dtdf %>% group_by(State) %>% summarise(Median_IBU = median(IBU)) 
ggplot(data = df2, aes(x= State, y = Median_IBU ,fill = Median_IBU))+geom_bar(stat='identity', width=0.7)+ ggtitle ("State v.Median_IBU") + scale_fill_gradient2(midpoint=mean(df2$Median_IBU), low= '#971B1E', mid = '#E02124', high = '#FFF200', space='#E1E1E1') + coord_flip() + geom_text(data = df2, aes(x = State, label = paste(State, Median_IBU, sep=' - ')), position = position_dodge(width = 1), hjust = -0.1, size = 4) + scale_y_continuous(limits = c(0,max(df2$Median_IBU) *1.1))

```

## Question 5: Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer? 
### As you can see from the barplots, we've found that Maine has the highest ABV at 6.7%  and the most bitter beer with an IBU of 61.

```{r "Question 5", message = FALSE, warning = FALSE}
#row number with of max value can be obtained by which.max, to get all information using by subsetting data frame. 

df1[which.max(df1$Median_ABV),] ###Maine
#IF we need to show the single highest ABV and IBU beers
#tail(arrange(dtdf, ABV))

df2[which.max(df2$Median_IBU),]  ###Maine

```

## Question 6: Comment on the summary statistics and distribution of the ABV variable.
### In analyzing the alcohol content of the craft beers, we see a roughly normal distribution along the median with the mean, represented in red, and median, in yellow, falling close to each other.  We do some outliers such as the max ABV of 12.5, more than double the mean.  On the low side, we as well see a small number with a low ABV.

```{r "Question 6", message = FALSE, warning = FALSE}
summary(dtdf$ABV)

sd(dtdf$ABV*100)   ###0.005437061
mean(dtdf$ABV*100)  ### 0.05564

ggplot(dtdf, aes(x = ABV*100)) + geom_histogram(aes(y = ..count..), fill = '#971B1E') + scale_y_continuous(name = "Frequency") + theme(legend.position = 'none') + scale_x_continuous(name = 'Median ABV (%)', breaks = seq(2,13,.5)) + geom_vline(aes(xintercept = mean(dtdf$ABV *100)),col='#FFF200',size=1) + geom_vline(aes(xintercept = median(dtdf$ABV *100)),col='#E02124',size=1)


```

## Question 7: Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
### While there appears to be some positive relationship between ABV and IBU, it appears to be somewhat weak.  There are numerous outliers, specifically those with a high ABV but low IBU, that don't fit well within that explanation.  This isn't unexpected upon additional review of the data, as seen below, as IPAs are known for their bitterness and higher alcohol content.  Taking a look at what type of beers have the highest ABV while the lowest IBU shows that that group consists primarily with beers you would expect to find lower IBUs such as Ales and Witbeer.
```{r "Question 7", message = FALSE, warning = FALSE}

dtdf %>% ggplot(aes(x=ABV,y= IBU)) + geom_point() + geom_smooth (formula = y~ x,method="lm",col='#E02124',size=1) + ggtitle(("Relationship between ABV and IBU"))
outliers <- dtdf[dtdf$ABV > 0.075, ]
outliers <- outliers[outliers$IBU < 25, ]
outliers['Style']

```

### Question 8: Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually.

```{r "Question 8", message = FALSE, warning = FALSE}
library(dplyr)
library(class)
library(caret)
library(stringr)

tdf <- merge(BeerN,Breweries, by = "Brew_ID", na.rm = True) %>% rename(Beer = Name.x, Brewery = Name.y)

TD <- tdf %>% filter(grepl('IPA|Ale',Beer,ignore.case = T)) 
 
TD = TD[!is.na(TD$ABV),]

TD= TD[!is.na(TD$IBU), ]

#dim(TD)
#get just the 3 relevant columns and get rid of the rest
TDSubset <- select(TD, ABV, IBU, Style)
#make style a character from factor
TDSubset$Style <- as.character(TDSubset$Style)

#make all IPA's have Style of IPA and all ales have simply ale - eliminates all of the additional types to act on them as all the same In response to Novin'ssuggestion to Ranjan, changing to 0 = ipa and 1 = ale
for (i in 1:nrow(TDSubset)) {
  if (is.na(str_match(TDSubset[i, 3], 'Ale'))) {
    TDSubset[i,3] <- 0 
    }
  else { 
    TDSubset[i,3] <- 1
    }
  }

head(TDSubset)
#take 70% of the TDSubset to use as train and the remainder as test
ind = sample((1:nrow(TDSubset)),.7*nrow(TDSubset))
train = TDSubset[ind,]
test = TDSubset[-ind,]

pred = c("ABV","IBU")

#run 50 iterations to test the best value of K to test the best accuracy
bestKResults <- data.frame(Acc = numeric(100), k = numeric(100))
for (i in 1:50) {
  class <- knn(train[,pred],test[,pred],train$Style , prob= TRUE, k= i)
  conMatrix <- confusionMatrix(table(class, test$Style))
  bestKResults$Acc[i] <- conMatrix$overall[1]
  bestKResults$k[i] <- i
  
}

ggplot(bestKResults, aes(x=k, y = Acc)) + geom_line()+ labs(x = "K Value", y = "Accuracy")


classifications = knn(train[,pred],test[,pred],train$Style , prob= TRUE, k= 17)
confusionMatrix(table(classifications, test$Style))

